When does vapor pressure deficit drive or reduce evapotranspiration?

### 2018-12-17 doc/paper/hess-revision/review-1-pierre-comments.tex
added pierre's revision comments

Just a quick general comment before addressing your specific
comments. By including a lot of content on PFT-focused analysis we
failed to communicate the major goal and scope of our project: we are
trying to characterize the response of ET to VPD, with all other
environment variables held fixed. To accomplish this, we need to
formulate an explicit function of ET in terms of environmental
variables and parameters, where any parameters can be approximated as
constant with regards to some [arbitrary] VPD perturbation
scenario. We will elaborate more below on why we used PFT-focused
scenarios for much of our analysis. However, the goal of our
manuscript was much more simple and fundamental: at a given place or
time, if you introduce a perturbation to VPD, what is the immediate ET
response (e.g. positive or negative)? Answering this question does not
require the much stronger assertion that any parameters must be
invariant within a given PFT.

@@

We agree that the uncertainties are large both within a PFT and across
a PFT. Looking back, we believe that some of our language
communicating \cite{Lin_2018}, \cite{Medlyn_2017}, and
\cite{Zhou_2014}'s results was misleading, and this was exacerbated by
our focus on PFT-analysis. We will remove that language from the
manuscript, and add language to better communicate our results as
reflections of the considerable within-PFT uncertainty.

We think comparing the rankings of given PFT values for g$_1$ and uWUE
can be misleading, given the magnitude of the uncertainties
involved. Most of our calculated values for uWUE are within one
standard deviation of \cite{Zhou_2015}'s results, but these deviations
can result in some changes in the ordering of PFTs from high to low
uWUE. The bigger problem in our eyes is that we made a mistake with
some language suggesting within intra-PFT variability is less than
inter-PFT variability, but clearly this is not the case. We will
remove this language. This misinterpretation should not have made the
manuscript; we actually included \cite{Zhou_2015}'s results in Table 2
to be transparent about the within-PFT uncertainties. g$_1$ values
also exhibit considerable uncertainty, and again, while the relative
rankings of these values may change, they are all within observed
ranges in \cite{Medlyn_2017}. We did not include these in Table 2
because \cite{Medlyn_2017} does not provide the numerical values;
however, we think it would be useful to provide estimates of these
ranges from \cite{Medlyn_2017}'s figures in our Table 2.

Regarding how to interpret g$_1$ and uWUE: we think the most best way
is with established physical relations from \cite{MEDLYN_2011}
(Equation 3 in the manuscript) and \cite{Zhou_2014} (Pg. 7, line 10 in
the manuscript), with the knowledge that there is uncertainty
involved. The quantities in these relationships all have some
intrinsic physical meaning, but can vary substantially within a
PFT. Again, we need to alter our language to better reflect this
intra-PFT variability.

However, \cite{Zhou_2014} did establish a constant uWUE approximation
as a good approximation for capturing the relationship between GPP, ET
and VPD at a given place and time, so it is still a very useful and
robust approximation for answering our research question (see
discussion earlier in this comment). Additionally, a constant g$_1$
approximation is used in many earth system models, so while it
introduces considerable uncertainty, it also makes our framework
useful for interpreting modeled vegetation response in ESMs and GCMs.

@@

What we need to communicate better is that a given ecosystem's state
at a place or time controls its response to a VPD perturbation. We are
making an approximation that we can parameterize the effect of the
ecoystem's state on VPD response with g$_1$ and uWUE. These quantities
can also vary due to soil moisture condition; the approximation we
need to make to answer our research question is that they are fixed
with respect to a VPD perturbation. In this sense, we do not view the
two statements: ``plants that are evolved to bred to prioritize
primary production over water conservation (e.g., crops) exhibit a
higher likelihood of atmospheric demand-driven response'' and
``ecosystem types are responding in this way because they have, on the
whole, been subjected to less soil water limitation (due to the
non-negligible effects of irrigation?)'' as mutually exclusive. In
fact, we view them as consistent with a view that crops and their
physiology (parameterized by g$_1$ and uWUE) exist at a given time and
place because they have not been subjected to soil water limitation,
and they have a given response to VPD because of their physiology
(which is a direct effect of the environment). In this way, climate
and land surface state are causes of the VPD response both directly
and through their effect on plant physiology (parameterized by g$_1$
and uWUE).

When writing this manuscript, there was definitely some internal
tension and debate about how to frame the analysis and results. We
could either focus on PFT-oriented results as previous literature has
done (e.g. holding plant physiology fixed within a given PFT), or
allow plant physiology terms to vary through time and space and look
at the distribution of ET response to VPD (see Section 3.5). After
re-examining the manuscript both after some time away from the
problem and in light of your comments we think a strong argument could
be made that we made the wrong choice with respect to this focus. It
may have made more sense to focus our analysis on the ET response more
generally across space and time as ecosystem-scale plant physiology
varies in response to climate and soil moisture.

@@

In order to fix the manuscript and our communication of the answer to
the question ``When does VPD drive or reduce ET?'', we see a few
potential paths that we will consider between now and the final
response after the discussion period. If you (or anyone else) has any
opinions or comments in the meantime, we would appreciate the insight
and feedback.

\begin{enumerate}
\item We include the discussion presented in this review and response,
  and include language explicitly stating that the purpose of the PFT
  analysis is to provide connections to other PFT-constant analyses
  and models (e.g. ESMs and GCMs). We will add extensive language on
  the uncertainty and weaknesses of this approach, and reframe our
  conclusions to reflect his uncertainty.
\item We instead alter our analysis to look at how VPD response varies
  with climate and general plant physiological variability, instead of
  focussing only on PFT analysis that poorly captures all of the observed
  variability in ET response. Sections 3.1 - 3.2 would be replaced by
  analysis direcly relating ET response to general climate and plant
  physiological terms, rather than PFT-mean analysis. For example, the
  ``scaling term'' analysis would be presented in terms of generic
  changes in plant height and temperature, and the ``sign term''
  analysis would be framed by generic changes to g$_1$ and uWUE, as
  informed by previous literature.
\item This is the most extreme option in terms of modifying the
  manuscript. We significantly alter the manuscript and instead focus
  on just the general shape of the ET-VPD curve (with environmental
  variables held fixed). This was not discussed in the review, but one
  of our most noteworthy results is on the general shape of the ET-VPD
  curve being concave up, given an assumption of a square-root VPD
  dependence of ET. This result is independent of any assumptions of
  uWUE and g$_1$, and to our knowledge it is first derived curve of
  ecosystem response to VPD (see Section 3.7). It also highlights the
  importance of discerning the exact exponent of VPD dependence, as
  this alters the fundamental nature and shape of the
  curve. Essentially, the manuscript would become just our motivation
  and derivation, followed by Section 3.7 exploring the consequences
  of the derivation for the shape of the VPD curve. This alteration of
  the manuscript would likely result in more of a technical note-type
  paper, and we are hesitant to do this because we think there is
  still a lot of useful information obtained by tying our results to
  real-world scenarios (as in Options 1 and 2 above).
\end{enumerate}

>>>>

Just a quick general comment before addressing your specific
comments. By including a PFT-focused analysis we did not fully communicate the major goal and scope of our project: we are
trying to characterize the response of ET to VPD, with all other
environment variables held fixed. To accomplish this, we need to
formulate an explicit function of ET in terms of environmental
variables and parameters, where any parameters can be approximated as
constant with regards to some [arbitrary] VPD perturbation
scenario. We will elaborate more below on why we used PFT-focused
scenarios for much of our analysis. However, the goal of our
manuscript was much more simple and fundamental: at a given place or
time, if you introduce a perturbation to VPD, what is the immediate ET
response (e.g. positive or negative)? Answering this question does not
require the much stronger assertion that any parameters must be
invariant within a given PFT.

@@

We agree that the uncertainties are large both within a PFT and across
PFTs. Looking back, we believe that some of our language
communicating \cite{Lin_2018}, \cite{Medlyn_2017}, and
\cite{Zhou_2014}'s results was misleading, and this was exacerbated by
our focus on PFT-analysis. We will remove that language from the
manuscript, and add language to better communicate our results as
reflections of the considerable within-PFT uncertainty.

We think comparing the rankings of given PFT values for g$_1$ and uWUE
can be misleading, given the magnitude of the uncertainties
involved. Most of our calculated values for uWUE are within one
standard deviation of \cite{Zhou_2015}'s results, but these deviations
can result in some changes in the ordering of PFTs from high to low
uWUE. The bigger problem in our eyes is that we made a mistake with
some language suggesting within intra-PFT variability is less than
inter-PFT variability, but clearly this is not the case. We will
remove this language. This misinterpretation should not have made the
manuscript SOMETHING MISSING HERE; we actually included \cite{Zhou_2015}'s results in Table 2
to be transparent about the within-PFT uncertainties. g$_1$ values
also exhibit considerable uncertainty, and again, while the relative
rankings of these values may change, they are all within observed
ranges in \cite{Medlyn_2017}. We did not include these in Table 2
because \cite{Medlyn_2017} does not provide the numerical values;
however, we think it would be useful to provide estimates of these
ranges from \cite{Medlyn_2017}'s figures in our Table 2.

Regarding how to interpret g$_1$ and uWUE: we think the best way
is with established physical relations from \cite{MEDLYN_2011}
(Equation 3 in the manuscript) and \cite{Zhou_2014} (Pg. 7, line 10 in
the manuscript), with the knowledge that there is uncertainty
involved. The quantities in these relationships all have some
intrinsic physical meaning, but can vary substantially within a
PFT. Again, we need to alter our language to better reflect this
intra-PFT variability.

However, \cite{Zhou_2014} did establish a constant uWUE approximation
as a good approximation for capturing the relationship between GPP, ET
and VPD at a given place and time, so it is still a very useful and
robust approximation for answering our research question (see
discussion earlier in this comment). Additionally, a constant g$_1$
approximation is used in many earth system models, so while it
introduces uncertainty, it also makes our framework
useful for interpreting modeled vegetation response in ESMs and GCMs.

@@

What we need to communicate better is that a given ecosystem's state
at a place or time controls its response to a VPD perturbation. We are
making an approximation that we can parameterize the effect of the
ecoystem's state on VPD response with g$_1$ and uWUE. These quantities
can also vary due to soil moisture condition; the approximation we
need to answer our research question is that they are fixed
with respect to a VPD perturbation. In this sense, we do not view the
two statements: ``plants that are evolved to bred to prioritize
primary production over water conservation (e.g., crops) exhibit a
higher likelihood of atmospheric demand-driven response'' and
``ecosystem types are responding in this way because they have, on the
whole, been subjected to less soil water limitation (due to the
non-negligible effects of irrigation?)'' as mutually exclusive. In
fact, we view them as consistent with a view that crops and their
physiology (parameterized by g$_1$ and uWUE) exist at a given time and
place because they have not been subjected to soil water limitation,
and they have a given response to VPD because of their physiology
(which is a direct effect of the environment). In this way, climate
and land surface state are causes of the VPD response both directly
and through their effect on plant physiology (parameterized by g$_1$
and uWUE).

When writing this manuscript, there was definitely some discussion as to how to best frame the analysis and results. We
could either focus on PFT-oriented results as previous literature has
done (e.g. holding plant physiology fixed within a given PFT), or
allow plant physiology terms to vary through time and space and look
at the distribution of ET response to VPD (see Section 3.5). After
re-examining the manuscript both after some time away from the
problem and in light of your comments we think a strong argument could
be made that we made the wrong choice with respect to this focus. It
may have made more sense to focus our analysis on the ET response more
generally across space and time as ecosystem-scale plant physiology
varies in response to climate and soil moisture.

@@

In order to improve the manuscript and our communication of the answer to
the question ``When does VPD drive or reduce ET?'', we see a few
potential paths that we will consider between now and the final
response after the discussion period. If you (or anyone else) has any
opinions or comments in the meantime, we would appreciate the insight
and feedback.

\begin{enumerate}
\item We include the discussion presented in this review and response,
  and include language explicitly stating that the purpose of the PFT
  analysis is to provide connections to other PFT-constant analyses
  and models (e.g. ESMs and GCMs). We will add extensive language on
  the uncertainty and weaknesses of this approach, and reframe our
  conclusions to reflect his uncertainty.
\item We instead alter our analysis to look at how VPD response varies
  with climate and general plant physiological variability, instead of
  focussing only on PFT analysis that poorly captures all of the observed
  variability in ET response. Sections 3.1 - 3.2 would be replaced by
  analysis directly relating ET response to general climate and plant
  physiological terms, rather than PFT-mean analysis. For example, the
  ``scaling term'' analysis would be presented in terms of generic
  changes in plant height and temperature, and the ``sign term''
  analysis would be framed by generic changes to g$_1$ and uWUE, as
  informed by previous literature.
\item This is the most extreme option in terms of modifying the
  manuscript. We significantly alter the manuscript and instead focus
  on just the general shape of the ET-VPD curve (with environmental
  variables held fixed). This was not discussed in the review, but one
  of our most noteworthy results is on the general shape of the ET-VPD
  curve being concave up, given an assumption of a square-root VPD
  dependence of ET. This result is independent of any assumptions of
  uWUE and g$_1$, and to our knowledge it is first derived curve of
  ecosystem response to VPD (see Section 3.7). It also highlights the
  importance of discerning the exact exponent of VPD dependence, as
  this alters the fundamental nature and shape of the
  curve. Essentially, the manuscript would become just our motivation
  and derivation, followed by Section 3.7 exploring the consequences
  of the derivation for the shape of the VPD curve. This alteration of
  the manuscript would likely result in more of a technical note-type
  paper, and we are hesitant to do this because we think there is
  still a lot of useful information obtained by tying our results to
  real-world scenarios (as in Options 1 and 2 above).
\end{enumerate}
